\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-08-21 paper/old/main-old.tex

\documentclass[preprint,12pt,a4paper]{elsarticle}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage{float}
\restylefloat{table}
\usepackage[colorlinks=true, urlcolor=blue, pdfborder={0 0 0}]{hyperref}
\usepackage{breakurl}
\usepackage{placeins}
\usepackage{geometry}

\geometry{left=1in, right=1in, top=1in, bottom=1in, headsep=0pt}

@@

\title{Davos: a Python package ``smuggler'' for constructing
  lightweight reproducible notebooks}
\author{Paxton C. Fitzpatrick}
\author{Jeremy R. Manning\corref{cor}}
\ead{Jeremy.R.Manning@Dartmouth.edu}
\cortext[cor]{Corresponding author}
\address{Department of Psychological and Brain Sciences\\Dartmouth College, Hanover, NH 03755}

\begin{abstract}
  Reproducibility is a core requirement of modern scientific research.
  For computational research, reproducibility means that code should
  produce the same results, even when run on different systems.  A
  standard approach to ensuring reproducibility entails packaging a
  project's dependencies along with its primary code base.  Existing
  solutions vary in how deeply these dependencies are specified,
  ranging from virtual environments, to containers, to virtual
  machines.  Each of these existing solutions requires installing or
  setting up a system for running the desired code, increasing the
  complexity and time cost of sharing or engaging with reproducible
  science. Here, we propose a lighter-weight solution: the
  Davos package.  When used in combination with a
  notebook-based Python project, Davos provides a mechanism
  for specifying the correct versions of the project's
  dependencies directly within the code that requires them,
  and automatically installing them in an isolated environment
  when the code is run. The Davos package further
  ensures that these packages and specific versions are used every
  time the notebook's code is executed.  This enables researchers to
  share a complete reproducible copy of their code within a single
  Jupyter notebook file.
\end{abstract}

@@

\begin{table}[H]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Metadata value} \\
\hline
C1 & Current code version &  v0.2.0 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/ContextLab/davos/tree/v0.2.0} \\
\hline
C3 & Code Ocean compute capsule & \\
\hline
C4 & Legal Code License & MIT \\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, JavaScript, PyPI/pip, IPython, Jupyter, ipykernel, PyZMQ.\newline Additional tools used for tests: pytest, Selenium, Requests, mypy, GitHub Actions \\
\hline
C7 & Compilation requirements, operating environments, and
     dependencies & Dependencies:~Python $\geq 3.6$, packaging, setuptools.\newline Supported OSes: MacOS, Linux, Unix-like.\newline Supported IPython environments: Jupyter Notebooks, JupyterLab, Google Colaboratory, Binder, IDE-based notebook editors. \\
\hline
C8 & Link to developer documentation/manual & \url{https://github.com/ContextLab/davos\#readme} \\
\hline
C9 & Support email for questions & \href{mailto:contextualdynamics@gmail.com}{contextualdynamics@gmail.com} \\
\hline
\end{tabular}
\caption{Code metadata}
\label{}
\end{table}

@@

Within the Python~\cite{vanR95} community, external packages that are
published in the most popular repositories~\cite{Pyth03, cond15} are
associated with version numbers and tags that allow users to guarantee
they are installing exactly the same code across different computing
environments~\cite{CoghStuf13}.  While it is \textit{possible} to
manually install the intended version of every dependency of a Python
script or package, manually tracking down those dependencies can
impose a substantial burden on the user and create room for mistakes
and inconsistencies. Further, when dependency versions are left
unspecified, replicating the original computing environment becomes
difficult or impossible~\citep{PimeEtal19}.

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/shareable_code}
\caption{\small \textbf{Systems for sharing code within the Python
  ecosystem.}  From left to right: plain-text \textbf{Python
  scripts} (\texttt{.py} files) provide the most basic ``system''
  for sharing raw code.  Scripts may reference external packages, but
  those packages must be manually installed on other users' systems.
  Further, any checking needed to verify that the correct versions of
  those packages were installed must also be performed manually.
  \textbf{Jupyter notebooks} (\texttt{.ipynb} files) comprise embedded
  text, executable code, and media (including rendered figures, code
  output, etc.).  When the \textbf{Davos package} is imported
  into a Jupyter notebook, the notebook's functionality is extended to
  automatically install any required external packages (at their
  correct versions, when specified).  \textbf{Virtual environments}
  allow users to install an isolated copy of Python and all required
  dependencies. This typically entails distributing a configuration
  file (e.g., a \texttt{pyproject.toml}~\cite{CannEtal16} or
  \texttt{environment.yml} file) that specifies all project
  dependencies (including version numbers of external packages)
  alongside the primary code base. Users can then install a
  third-party tool~\cite[e.g.,][]{Anac12, Eust19} to read the file and
  build the environment.  \textbf{Containers} provide a means of
  defining an isolated environment that includes a complete operating
  system (independent of the user's operating system), in addition to
  (optionally) specifying a virtual environment or other
  configurations needed to provide the necessary computing
  environment.  Containers are typically defined using specification
  files (e.g., a plain-text \texttt{Dockerfile}) that instruct the
  virtualization engine regarding how to build the containerized
  environment.  \textbf{Virtual machines} provide a complete
  hardware-level simulation of the computing environment.  In addition
  to simulating specific hardware, virtual machines (typically
  specified using binary image files) must also define operating
  system-level properties of the computing environment.  Systems to
  the left of the blue vertical line entail sharing individual files,
  with no additional installation or configuration needed to run the
  target code.  Systems to the right of the red vertical line support
  precise control over dependencies and versioning.  Notebooks
  enhanced using the Davos package are easily shareable and
  require minimal setup costs, while also facilitating high
  reproducibility by enabling precise control over project
  dependencies.}
\label{fig:code-sharing}
\end{figure}

@@

We designed Davos to occupy a ``sweet spot'' between these extremes.
Davos is a notebook-installable package that adds functionality to the
default notebook experience. Like standard Jupyter notebooks,
Davos-enhanced notebooks allow researchers to include text, executable
code, and media within a single file. No further setup or installation is
required from the user, beyond what is needed to run standard Jupyter
notebooks. And like virtual environments, Davos provides a convenient
mechanism for fully specifying (and installing, as needed) a complete set of
Python dependencies, including specific package versions, which are contained
and isolated from the rest of the user's system.

@@

The Davos package is named after Davos Seaworth, a smuggler referred
to as ``the Onion Knight" from the series \textit{A Song of Ice and Fire} by
George R. R. Martin~\cite{Mart98}. The \texttt{smuggle} keyword provided by
Davos is a play on Python's \texttt{import} keyword: whereas importing
can load a package into the Python workspace within the existing rules and
frameworks provided by the Python language, ``smuggling'' provides an
alternative that expands the scope and reach of ``importing.'' Like the
character Davos Seaworth (who became famous for smuggling onions through a
blockade on his homeland), we use ``onion'' comments to precisely control how
packages are smuggled into the Python workspace.

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/package_structure}
\caption{\small \textbf{Package structure.} The Davos package
  comprises two interdependent subpackages.  The \texttt{davos.core}
  subpackage includes modules for parsing \texttt{smuggle} statements
  and onion comments, installing and validating packages, isolating and managing and
  configuring Davos's behavior.  The
  \texttt{davos.implementations} subpackage includes
  environment-specific modifications and features that are needed to
  support the core functionality across different notebook-based
  environments.  Individual modules (i.e., \texttt{.py} files) are represented by lime
  rounded rectangles, and arrows denote dependencies (each arrow
  points to a module that imports objects defined in the module at the
  arrow's source).}
\label{fig:package-structure}
\end{figure}

@@

The Davos package consists of two interdependent subpackages
(see Fig.~\ref{fig:package-structure}). The first,
\texttt{davos.core}, comprises a set of modules that
implement the bulk of the package's core
functionality, including pipelines for installing and validating
packages, custom parsers for the \texttt{smuggle} statement (see
Sec.~\ref{subsec:smuggle}) and onion comment (see
Sec.~\ref{subsec:onion}), a system for isolating dependencies of
different projects (see Sec.~\ref{subsec:projects}), and a runtime
interface for configuring Davos's behavior (see Sec.~\ref{subsec:config}).
However, certain critical aspects of this
functionality require (often substantially) different implementations
depending on properties of the notebook environment in which
Davos is used (e.g., whether the frontend is provided by
Jupyter or Google Colaboratory, or which version of
IPython~\cite{PereGran07} is used by the notebook kernel). To deal
with this, environment-dependent components of core features and behaviors
are isolated and abstracted to ``helper functions'' in the
\texttt{davos.implementations} subpackage. This second subpackage
defines multiple, interchangeable versions of each helper function,
organized into modules by the conditions that trigger their use. At
runtime, Davos detects various features in the notebook
environment and selectively imports a single version of each helper
function into the top-level \texttt{davos.implementations} namespace,
allowing \texttt{davos.core} modules to access the proper
implementations for the current notebook environment in a single,
consistent location. An additional benefit of this design is that it
allows both maintainers and users to easily extend Davos to
support new, updated, or custom notebook variants by adding new
\texttt{davos.implementations} modules that define their own versions
of each helper function, modified from existing implementations as
needed.

@@

Functionally, importing Davos in an IPython notebook enables
an additional Python keyword: ``\texttt{smuggle}'' (see
Sec.~\ref{subsec:implementation} for details on how this works).
The \texttt{smuggle} keyword can be used as a drop-in
replacement for Python's built-in \texttt{import} keyword to load
packages, modules, and other objects into the notebook's namespace.
However, whereas \texttt{import} will fail if the requested package is
not installed locally, \texttt{smuggle} statements can handle missing
packages on the fly.  If a smuggled package does not exist in the
user's Python environment, Davos will download and install it automatically,
expose its contents to Python's \texttt{import} machinery, and load it
into the notebook for immediate use.

Importantly, packages installed by Davos are made available for use in the
notebook without affecting the user's Python environment or existing packages.
By default, \texttt{smuggle} statements will install missing packages (and any
missing dependencies of those packages) into a notebook-specific, virtual
environment-like directory called a ``project'' (see
Sec.~\ref{subsec:projects}). In turn, \texttt{smuggle} statements executed in a
particular notebook will preferentially load packages from that notebook's
project directory whenever they are available, rather than searching for them
in the user's main Python environment. In this way, \texttt{smuggle}
statements can be substituted for \texttt{import} statements to automatically
ensure that all packages needed to run a notebook are installed and available
at runtime each time the notebook is run, without risking interfering with
dependencies of the user's other Python programs, or other Davos-enhanced
notebooks.

@@

For greater control over the behavior of \texttt{smuggle} statements, Davos
defines an additional construct called the ``onion comment.'' An onion comment
is a special type of inline comment that may be placed on a line containing a
\texttt{smuggle} statement to customize how Davos searches for the smuggled
package locally and, if necessary, downloads and installs it. Onion comments
follow a simple format based on the ``type comment'' syntax introduced in PEP
484~\cite{vanREtal14}, and are designed to make managing packages with Davos
intuitive and familiar. To construct an onion comment, users provide the name
of the installer program (e.g., \texttt{pip}) and the same arguments one would
use to manually install the package as desired via the command line:
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/snippet1}
\end{center}
Occasionally, a package's distribution name (i.e., the name used
when installing it) may differ from its top-level module name (i.e., the name
used when importing it). In such cases, an onion comment may be used to ensure
that Davos installs the proper package if it cannot be found locally:
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/snippet2}
\end{center}
Because onion comments may be constructed to specify any aspect of
the installer program's behavior, they provide a mechanism for precisely controlling
how, where, and when smuggled packages are installed. Critically, if an onion
comment includes a version specifier~\cite{CoghStuf13}, Davos will ensure that
the version of the package loaded into the notebook matches the specific
version requested, or satisfies the given version constraints. If the smuggled
package exists locally, Davos will extract its version information from its
metadata and compare it to the specifier provided. If the two are incompatible
(or no local installation is found), Davos will download, install, and load a
suitable version of the package instead:
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/snippet3}
\end{center}
Onion comments can also be used to \texttt{smuggle} specific VCS references (e.g.,
Git~\cite{TorvHama05} branches, commits, tags, etc.):
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/snippet4}
\end{center}
Davos processes onion comments internally before forwarding arguments to the
installer program. In addition to preventing shared notebooks from executing
arbitrary code in a user's shell, this enables Davos to adapt its behavior
based on how particular flags will affect the behavior of the installer
program. For example, including \texttt{pip}'s \texttt{--no-input} flag will also
temporarily enable Davos's non-interactive mode (see Sec.~\ref{subsec:config}).
Similarly, if an onion comment contains either the \texttt{-I}/\texttt{--ignore-installed},
\texttt{-U}/\texttt{--upgrade}, or \texttt{--force-reinstall} flag, Davos will
skip checking for a local copy of the smuggled package before installing a
new one:
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/snippet5}
\end{center}
Since the purpose of an onion comment is to describe how a smuggled package
should be installed (if necessary) so that it can be loaded and used
immediately, options that would cause the package not to be installed (such as
\texttt{-h}/\texttt{--help} or \texttt{--dry-run}) are disallowed. Additionally,
when using a Davos project to isolate smuggled packages (the default behavior;
see Sec.~\ref{subsec:projects}), onion comments may not contain options that
would change the package's installation location (such as
\texttt{-t}/\texttt{--target}, \texttt{--root}, or \texttt{--prefix}). However, if
the user disables project-based isolation and specifies \texttt{--target <dir>},
Davos will ensure that \texttt{<dir>} is included in the module search path (i.e.,
\texttt{sys.path}), prepending it if necessary, so the package can be loaded.

\subsubsection{Projects}\label{subsec:projects}

Standard approaches to installing packages from within a notebook can alter the local Python environment in potentially unexpected and undesired ways. For example, running a notebook that installs its dependencies via system shell commands (prefixed with ``\texttt{!}'') or IPython magic commands (prefixed with ``\texttt{\%}'') may cause other existing packages in the user's environment to be uninstalled and replaced with alternate versions. This can lead to incompatibilities between installed packages, affect the behavior of the user's other scripts or notebooks, or even interfere with system applications.

To prevent Davos-enhanced notebooks from having unwanted side-effects on the user's environment, Davos automatically isolates packages installed via \texttt{smuggle} statements using a custom scheme called ``projects.'' Functionally, a Davos project is similar to a standard Python virtual environment (e.g., created with the standard library's \texttt{venv} module or a third-party tool like \texttt{virtualenv}~\cite{BickEtal07}): it consists of a directory (within a hidden \texttt{.davos} folder in the user's home directory) that houses third-party packages needed for a particular project or task. However, Davos projects do not need to be manually activated and deactivated, do not contain separate Python or \texttt{pip} executables, and \textit{extend} the user's main Python environment rather than replace it.

When Davos is imported into a notebook, a notebook-specific project directory is automatically created (if it does not exist already). 
%When Davos is imported into a notebook, a notebook-specific project directory is automatically created (if it does not exist already), named for the absolute path to the notebook file.

Notebook-specific projects are named for the absolute path to the notebook file.

 %Davos projects function similarly to simplified versions of standard Python virtual environments (e.g., created with the standard library's \texttt{venv} module or a third-party tool like \texttt{virtualenv}~\cite{BickEtal07}) with a few differences: they do not need to be manually activated and deactivated, they do not contain separate Python or \texttt{pip} executables, and they \textit{extend} the main Python environment rather than replace it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADD THIS EITHER TO START OF PROJECTS SUBSUBSECTION OR IN IMPACT SECTION   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%A common way of avoiding this is to create a virtual environment in which to run the notebook, and instruct anyone with whom the notebook is shared to do the same. While effective, this added requirement introduces additional 
%
%
%(or group of related notebooks) that require different 
%
%
%\begin{itemize}
%  \item{common solution is to use a virtual environment}
%  \item{this introduces complexity}
%  \item{davos's solution is projects}
%\end{itemize}

\bigskip\bigskip\textcolor{red}{\textbf{========== TODO: finish editing from here to end ==========}}\bigskip\bigskip

%Because Davos can install new packages, running the code in a
%Davos-enhanced notebook might (in principle) affect the behavior of
%\textit{other} Python-based software (e.g., other notebooks, scripts, etc.) by
%altering which packages are installed in the runtime environment. This could
%lead to undesired consequences. For example, suppose Person A develops a
%notebook (Notebook A) for their research project. We will assume that Notebook
%A does not use Davos to manage project dependencies.  If Person A
%runs a Davos-enhanced Notebook B, e.g., sent by another developer,
%might this unexpectedly affect the behavior of Notebook A?

%To prevent unwanted
%changes to the user's Python environment, Davos incorporates its
%own virtual environment-like scheme for isolating packages it installs. When Davos
%is imported, a new virtual environment (folder) is created automatically. The
%folder's name may be customized to support multi-notebook projects. Any
%\texttt{smuggle}d packages that were not available in the notebook's runtime
%environment are installed to the current project folder. The runtime
%environment remains unaffected by Davos's behavior (see
%Sec.~\ref{subsec:projects}).
%By default, projects are notebook-specific, but can also be shared by multiple notebooks, and can be managed interactively from within a Davos-enhanced notebook.

%%%%%%%% JEREMY VERSION %%%%%%%%%
%Installing new packages in a notebook using standard approaches (e.g., system commands) affect the
%runtime environment. This could lead to undesired behaviors. For example, running a notebook that
%installs new packages in the user’s primary system environment might alter their main system installation
%and/or containerized environment in unexpected ways (e.g., changing package versions, causing conflicts
%with other packages, etc.). To protect against undesired changes to the runtime environment, Davos
%incorporates its own virtual environment for managing packages it installs. When Davos is imported,
%a new virtual environment (folder) is created automatically. The folder’s name may be customized to
%support multi-notebook projects. Any smuggled packages that were not available in the notebook’s
%runtime environment are installed to the current project folder. The runtime environment remains unaffected
%by Davos’s behavior (see Sec. 2.2.3).

We implemented a ``project'' system in Davos to protect against the
above scenario. By default, importing Davos creates a new project
folder in the user's home directory (contained within a hidden \texttt{.davos}
folder). The default project name is computed to uniquely identify each
notebook according to its filename and path. Any packages that were not
originally available in the notebook's runtime environment are installed to the
notebook's project directory. When external libraries are \texttt{smuggle}d,
Davos temporally appends the current project directory to the search
path. Because the user's system path remains unchanged, and because none of the
runtime environment's packages are altered, the user's system and runtime
environment remain unaffected (aside from installing the Davos package
itself to the runtime environment).

Each notebook's project may be customized by setting \texttt{davos.project} to
any string that can be used as a valid folder name in the user's operating
system. By customizing the project name, users can build multi-notebook
projects that share the same core set of dependencies without needing to
duplicate each package for each notebook in the project.

Finally, if the user \textit{does} wish to modify their runtime environment,
this may be done by setting \texttt{davos.project} to \texttt{None}. Doing so
will cause any packages installed by Davos to affect the user's
runtime environment.  This is generally not recommended, as it can lead to
unintended consequences for other code that shares the runtime environment.

\subsubsection{Configuring and querying Davos}\label{subsec:config}

Davos's behavior may be customized by modifying a set of attributes attached to
the \texttt{davos} module object that is added to the workspace when Davos is
imported. These attributes may be modified, displayed, or checked
programmatically at runtime (see Sec.~\ref{sec:illustrative-example} for an
illustrative example or Sec.~\ref{subsec:implementation} for implementation
details and additional information). These include:

\begin{itemize}
\item \texttt{.active}: This attribute controls whether support for \texttt{smuggle}
  statements and onion comments is enabled (\texttt{True}) or
  disabled (\texttt{False}).  When Davos is first imported,
  the \texttt{.active} attribute is set to \texttt{True}.

>>>>

\documentclass[preprint,12pt,a4paper]{elsarticle}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage{float}
\restylefloat{table}
\usepackage[colorlinks=true, urlcolor=blue, pdfborder={0 0 0}]{hyperref}
\usepackage{breakurl}
\usepackage{placeins}
\usepackage{geometry}
\geometry{left=1in,right=1in,top=1in,bottom=1in,headsep=0pt}

@@

\title{\texttt{davos}: a Python package ``smuggler'' for constructing
  lightweight reproducible notebooks}
\author{Paxton C. Fitzpatrick}
\author{Jeremy R. Manning\corref{cor}}
\ead{Jeremy.R.Manning@Dartmouth.edu}
\cortext[cor]{Corresponding author}
\address{Department of Psychological and Brain Sciences\\Dartmouth College, Hanover, NH 03755}

\begin{abstract}
  Reproducibility is a core requirement of modern scientific research.
  For computational research, reproducibility means that code should
  produce the same results, even when run on different systems.  A
  standard approach to ensuring reproducibility entails packaging a
  project's dependencies along with its primary code base.  Existing
  solutions vary in how deeply these dependencies are specified,
  ranging from virtual environments, to containers, to virtual
  machines.  Each of these existing solutions requires installing or
  setting up a system for running the desired code, increasing the
  complexity and time cost of sharing or engaging with reproducible
  science. Here, we propose a lighter-weight solution: the
  \texttt{davos} package.  When used in combination with a
  notebook-based Python project, \texttt{davos} provides a mechanism
  for specifying (and automatically installing) the correct versions
  of the project's dependencies.  The \texttt{davos} package further
  ensures that those packages and specific versions are used every
  time the notebook's code is executed.  This enables researchers to
  share a complete reproducible copy of their code within a single
  Jupyter notebook file.
\end{abstract}

@@

\begin{table}[H]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Metadata value} \\
\hline
C1 & Current code version &  v0.1.1 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/ContextLab/davos/tree/v0.1.1} \\
\hline
C3 & Code Ocean compute capsule & \\
\hline
C4 & Legal Code License & MIT \\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & Python, JavaScript, PyPI/pip, IPython, Jupyter, ipykernel, PyZMQ. Additional tools used for tests: pytest, Selenium, Requests, Mypy, GitHub Actions \\
\hline
C7 & Compilation requirements, operating environments, and
     dependencies & Dependencies:~Python $\geq 3.6$, packaging, setuptools.~Supported OSes: MacOS, Linux, Unix-like.~Supported IPython environments: Jupyter Notebooks, JupyterLab, Google Colaboratory, Binder, IDE-based notebook editors. \\
\hline
C8 & Link to developer documentation/manual & \url{https://github.com/ContextLab/davos\#readme} \\
\hline
C9 & Support email for questions & contextualdynamics@gmail.com \\
\hline
\end{tabular}
\caption{Code metadata}
\label{}
\end{table}

@@

Within the Python~\cite{vanR95} community, external packages that are
published in the most popular repositories~\cite{Pyth03, cond15} are
associated with version numbers and tags that allow users to guarantee
they are installing exactly the same code across different computing
environments~\cite{CoghStuf13}.  While it is \textit{possible} to
manually install the intended version of every dependency of a Python
script or package, manually tracking down those dependencies can
impose a substantial burden on the user and create room for mistakes
and inconsistencies. Further, when dependency versions are left
unspecified, replicating the original computing environment becomes
difficult or impossible.

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/shareable_code}
\caption{\small \textbf{Systems for sharing code within the Python
    ecosystem.}  From left to right: plain-text \textbf{Python
    scripts} (\texttt{.py} files) provide the most basic ``system''
  for sharing raw code.  Scripts may reference external packages, but
  those packages must be manually installed on other users' systems.
  Further, any checking needed to verify that the correct versions of
  those packages were installed must also be performed manually.
  \textbf{Jupyter notebooks} (\texttt{.ipynb} files) comprise embedded
  text, executable code, and media (including rendered figures, code
  output, etc.).  When the \textbf{\texttt{davos} package} is imported
  into a Jupyter notebook, the notebook's functionality is extended to
  automatically install any required external packages (at their
  correct versions, when specified).  \textbf{Virtual environments}
  allow users to install an isolated copy of Python and all required
  dependencies. This typically entails distributing a configuration
  file (e.g., a \texttt{pyproject.toml}~\cite{CannEtal16} or
  \texttt{environment.yml} file) that specifies all project
  dependencies (including version numbers of external packages)
  alongside the primary code base. Users can then install a
  third-party tool~\cite[e.g.,][]{Anac12, Eust19} to read the file and
  build the environment.  \textbf{Containers} provide a means of
  defining an isolated environment that includes a complete operating
  system (independent of the user's operating system), in addition to
  (optionally) specifying a virtual environment or other
  configurations needed to provide the necessary computing
  environment.  Containers are typically defined using specification
  files (e.g., a plain-text \texttt{Dockerfile}) that instruct the
  virtualization engine regarding how to build the containerized
  environment.  \textbf{Virtual machines} provide a complete
  hardware-level simulation of the computing environment.  In addition
  to simulating specific hardware, virtual machines (typically
  specified using binary image files) must also define operating
  system-level properties of the computing environment.  Systems to
  the left of the blue vertical line entail sharing individual files,
  with no additional installation or configuration needed to run the
  target code.  Systems to the right of the red vertical line support
  precise control over dependencies and versioning.  Notebooks
  enhanced using the \texttt{davos} package are easily shareable and
  require minimal setup costs, while also facilitating high
  reproducibility by enabling precise control over project
  dependencies.}
\label{fig:code-sharing}
\end{figure}

@@

We designed \texttt{davos} to occupy a ``sweet spot'' between these
extremes.  \texttt{davos} is a notebook-installable package that adds
functionality to the default notebook experience. Like standard
Jupyter notebooks, \texttt{davos}-enhanced notebooks allow researchers
to include text, executable code, and media within a single file. No
further setup or installation is required, beyond what is needed to
run standard Jupyter notebooks. And like virtual environments,
\texttt{davos} provides a convenient mechanism for fully specifying
(and installing, as needed) a complete set of Python dependencies,
including specific package versions.

@@

The \texttt{davos} package is named after Davos Seaworth, a smuggler
referred to as ``the Onion Knight" from the series \textit{A Song of
  Ice and Fire} by George R. R. Martin~\cite{Mart98}.  The
\texttt{smuggle} keyword provided by \texttt{davos}
is a play on Python's \texttt{import} keyword: whereas importing can
load a package into the Python workspace within the existing rules
and frameworks provided by the Python language, ``smuggling'' provides
an alternative that expands the scope and reach of ``importing.''
Like the character Davos Seaworth (who became famous for smuggling
onions through a blockade on his homeland), we use ``onion'' comments
to precisely control how packages are smuggled into the Python
workspace.

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/package_structure}
\caption{\small \textbf{Package structure.} The \texttt{davos} package
  comprises two interdependent subpackages.  The \texttt{davos.core}
  subpackage includes modules for parsing \texttt{smuggle} statements
  and onion comments, installing and validating packages, and
  configuring \texttt{davos}'s behavior.  The
  \texttt{davos.implementations} subpackage includes
  environment-specific modifications and features that are needed to
  support the core functionality across different notebook-based
  environments.  Individual modules (i.e., \texttt{.py} files) are represented by lime
  rounded rectangles, and arrows denote dependencies (each arrow
  points to a module that imports objects defined in the module at the
  arrow's source).}
\label{fig:package-structure}
\end{figure}

@@

The \texttt{davos} package consists of two interdependent subpackages
(see Fig.~\ref{fig:package-structure}). The first,
\texttt{davos.core}, comprises a set of modules that
implement the bulk of the package's core
functionality, including pipelines for installing and validating
packages, custom parsers for the \texttt{smuggle} statement (see
Sec.~\ref{subsec:smuggle}) and onion comment (see
Sec.~\ref{subsec:onion}), and a runtime interface for configuring
\texttt{davos}'s behavior (see Sec.~\ref{subsec:config}). However,
certain critical aspects of this
functionality require (often substantially) different implementations
depending on properties of the notebook environment in which
\texttt{davos} is used (e.g., whether the frontend is provided by
Jupyter or Google Colaboratory, or which version of
IPython~\cite{PereGran07} is used by the notebook kernel). To deal
with this, environment-dependent parts of core features and behaviors
are isolated and abstracted to ``helper functions'' in the
\texttt{davos.implementations} subpackage. This second subpackage
defines multiple, interchangeable versions of each helper function,
organized into modules by the conditions that trigger their use. At
runtime, \texttt{davos} detects various features in the notebook
environment and selectively imports a single version of each helper
function into the top-level \texttt{davos.implementations} namespace,
allowing \texttt{davos.core} modules to access the proper
implementations for the current notebook environment in a single,
consistent location. An additional benefit of this design is that it
allows maintainers, developers, and users to extend \texttt{davos} to
support new, updated, or custom notebook variants by creating new
\texttt{davos.implementations} modules that define their own versions
of each helper function, modified from existing implementations as
needed.

@@

Functionally, importing \texttt{davos} in an IPython notebook enables
an additional Python keyword: ``\texttt{smuggle}'' (see
Sec.~\ref{subsec:implementation} for details on how this works).
The \texttt{smuggle} keyword-like object can be used as a drop-in
replacement for Python's built-in \texttt{import} keyword to load
packages, modules, and other objects into the current namespace.
However, whereas \texttt{import} will fail if the requested package is
not installed locally, \texttt{smuggle} statements can handle missing
packages on the fly.  If a smuggled package does not exist in the
local environment, \texttt{davos} will download and install it automatically,
expose its contents to Python's \texttt{import} machinery, and load it
into the namespace for immediate use.

@@

For greater control over the behavior of \texttt{smuggle} statements,
\texttt{davos} defines an additional construct called the ``onion
comment.'' An onion comment is a special type of inline comment that
may be placed on a line containing a \texttt{smuggle} statement to
customize how \texttt{davos} searches for the smuggled package locally
and, if necessary, downloads and installs it. Onion comments follow a
simple format based on the ``type comment'' syntax introduced in PEP
484~\cite{vanREtal14}, and are designed to make managing packages with
\texttt{davos} intuitive and familiar. To construct an onion comment,
users provide the name of the installer program (e.g., \texttt{pip})
and the same arguments one would use to manually install the package
as desired via the command line:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet1}
\end{center}
Occasionally, a package's distribution name (i.e., the name used when
installing it) may differ from its top-level module name (i.e., the
name used when importing it). In such cases, an onion comment may be
used to ensure that \texttt{davos} installs the proper package if it
cannot be found locally:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet2}
\end{center}
Because onion comments may be constructed to specify any aspect of the
installer's behavior, they provide a mechanism for precisely
controlling how, where, and when smuggled packages are installed.
Critically, if an onion comment includes a version
specifier~\cite{CoghStuf13}, \texttt{davos} will ensure that the
version of the package loaded into the notebook matches the specific
version requested, or satisfies the given version constraints. If the
smuggled package exists locally, \texttt{davos} will extract its
version information from its metadata and compare it to the specifier
provided. If the two are incompatible (or no local installation is
found), \texttt{davos} will download, install, and load a suitable
version of the package instead:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet3}
\end{center}
Onion comments can also be used to \texttt{smuggle} specific VCS references
(e.g., Git~\cite{TorvHama05} branches, commits, tags, etc.):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet4}
\end{center}
\texttt{davos} processes onion comments internally before forwarding
arguments to the installer program. In addition to preventing onion
comments from being used as a vehicle for shell injection attacks,
this enables \texttt{davos} to adapt its behavior based on how particular
flags will affect the behavior of the installer program. For example, if an
onion comment contains either the \texttt{-I}/\texttt{--ignore-installed},
\texttt{-U}/\texttt{--upgrade}, or \texttt{--force-reinstall} flag,
\texttt{davos} will not bother checking for a local copy of the smuggled
package before installing a new one:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet5}
\end{center}
Similarly, the \texttt{--no-input} flag will temporarily enable
\texttt{davos}'s non-interactive mode (see
Sec.~\ref{subsec:config}), and installing a smuggled package into a
custom directory (\texttt{<dir>}) using the \texttt{--target <dir>}
flag will cause \texttt{davos} to prepend \texttt{<dir>} to the module
search path (i.e., \texttt{sys.path}), if necessary, so the package
can be imported.

\subsubsection{The \texttt{davos} config object}\label{subsec:config}

The \texttt{davos} config object provides a high-level interface for
controlling various aspects of \texttt{davos}'s behavior. After
importing \texttt{davos}, the \texttt{davos.con\-fig} object (a
singleton) exposes configurable options as attributes that can be
modified, displayed in the notebook, or checked programmatically at runtime
(see Sec.~\ref{sec:illustrative-example} for an illustrative example or
Sec.~\ref{subsec:implementation} for implementation details and
additional information). These include:

\begin{itemize}
\item \texttt{.active}: This attribute controls whether support for \texttt{smuggle}
  statements and onion comments is enabled (\texttt{True}) or
  disabled (\texttt{False}).  When \texttt{davos} is first imported,
  the \texttt{.active} attribute is set to \texttt{True}.

\item \texttt{.auto\_rerun}: This attribute controls how
  \texttt{davos} behaves when attempting to \texttt{smuggle} a new
  version of a package that was previously imported and cannot be
  reloaded. This can happen if the package includes extension modules
  that dynamically link C or C++ objects to the Python interpreter,
  and the code that generates those objects was changed between the
  previously imported and to-be-smuggled versions.  If this attribute
  is set to \texttt{True}, \texttt{davos} will automatically restart
  the notebook kernel and rerun all code up to (and including) the
  current \texttt{smuggle} statement. If set to \texttt{False} (the default),
  \texttt{davos} will instead issue a warning, pause execution, and
  prompt the user to either restart and rerun the notebook, or
  continue running with the previously imported package version until
  the next time the kernel is restarted manually.  Note that, as of
  this writing, the \texttt{.auto\_rerun} attribute is not supported
  in Google Colaboratory notebooks.

\item \texttt{.confirm\_install}: If set to \texttt{True} (default:
  \texttt{False}), \texttt{davos} will require user confirmation
  before installing a smuggled package that does not yet exist in the
  user's environment.

\item \texttt{.noninteractive}: Setting this attribute to
  \texttt{True} (default: \texttt{False}) enables non-in\-ter\-act\-ive
  mode, in which all user interactions (prompts and dialogues) are
  disabled. Note that in non-interactive mode, the
  \texttt{confirm\_install} option is set to \texttt{False}.  If
  \texttt{auto\_rerun} is set to \texttt{False} while in non-interactive
  mode, \texttt{davos} will raise an exception if a smuggled package
  cannot be reloaded, rather than prompting the user.
